---
title: "CDPH Weekly Surveillance Report"
subtitle: "California Disease Surveillance Summary"
author: "California Department of Public Health - Epidemiology Division"
date: today
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
execute:
  warning: false
  message: false
params:
  start_date: "2026-01-01"
  end_date: "2026-01-07"
  county: "All"
---

```{r setup, include=FALSE}
library(tidyverse)
library(gt)
library(scales)
library(plotly)

# Parameters
START_DATE <- as.Date(params$start_date)
END_DATE <- as.Date(params$end_date)
COUNTY_FILTER <- params$county

# Load data
surveillance <- read_csv("../data/disease_surveillance.csv", show_col_types = FALSE) |>
  mutate(report_date = as.Date(report_date))

outbreaks <- read_csv("../data/outbreak_summary.csv", show_col_types = FALSE)

# Filter to reporting period
report_data <- surveillance |>
  filter(report_date >= START_DATE, report_date <= END_DATE)

if (COUNTY_FILTER != "All") {
  report_data <- report_data |> filter(county == COUNTY_FILTER)
}

# Calculate summary metrics
total_cases <- nrow(report_data)
total_hosp <- sum(report_data$hospitalized)
total_deaths <- sum(report_data$outcome == "Deceased")
pct_lab_confirmed <- mean(report_data$lab_confirmed) * 100
avg_days_report <- mean(report_data$days_to_report, na.rm = TRUE)
```

**DISCLAIMER:** This report contains synthetic data for demonstration purposes only.

## Report Period

| Parameter | Value |
|-----------|-------|
| **Start Date** | `r format(START_DATE, "%B %d, %Y")` |
| **End Date** | `r format(END_DATE, "%B %d, %Y")` |
| **Geographic Scope** | `r COUNTY_FILTER` |
| **Generated** | `r format(Sys.time(), "%B %d, %Y at %H:%M")` |

## Executive Summary

```{r exec-summary, echo=FALSE, results='asis'}
# Determine alert level
if (total_deaths > 10 | total_hosp > 100) {
  alert_level <- "HIGH"
  alert_color <- "#DC143C"
} else if (total_hosp > 50) {
  alert_level <- "MODERATE"
  alert_color <- "#FF8C00"
} else {
  alert_level <- "ROUTINE"
  alert_color <- "#2E8B57"
}

cat(sprintf("### Alert Level: <span style='color: %s; font-weight: bold;'>%s</span>\n\n",
            alert_color, alert_level))
```

| Metric | Value |
|--------|-------|
| **Total Cases** | `r format(total_cases, big.mark = ",")` |
| **Hospitalizations** | `r format(total_hosp, big.mark = ",")` |
| **Deaths** | `r total_deaths` |
| **Lab Confirmation Rate** | `r sprintf("%.1f%%", pct_lab_confirmed)` |
| **Avg Days to Report** | `r sprintf("%.1f", avg_days_report)` |

## Cases by Condition

```{r condition-table, echo=FALSE}
condition_summary <- report_data |>
  group_by(condition) |>
  summarise(
    Cases = n(),
    Hospitalized = sum(hospitalized),
    `Hosp Rate` = sprintf("%.1f%%", sum(hospitalized) / n() * 100),
    Deaths = sum(outcome == "Deceased"),
    `Outbreak-Associated` = sum(is_outbreak_associated),
    .groups = "drop"
  ) |>
  arrange(desc(Cases))

condition_summary |>
  gt() |>
  tab_header(
    title = "Disease Condition Summary",
    subtitle = sprintf("%s to %s", format(START_DATE, "%b %d"), format(END_DATE, "%b %d, %Y"))
  ) |>
  cols_align(align = "right", columns = -condition) |>
  tab_options(table.font.size = px(12))
```

## Cases by County

```{r county-plot, echo=FALSE, fig.width=10, fig.height=6}
county_data <- report_data |>
  group_by(county) |>
  summarise(cases = n(), .groups = "drop") |>
  arrange(desc(cases)) |>
  head(15)

p <- ggplot(county_data, aes(x = reorder(county, cases), y = cases, fill = cases)) +
  geom_col() +
  coord_flip() +
  scale_fill_gradient(low = "#cce5ff", high = "#003366") +
  labs(
    title = "Top 15 Counties by Case Count",
    x = NULL,
    y = "Number of Cases"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

ggplotly(p)
```

## Cases by Age Group

```{r age-table, echo=FALSE}
age_summary <- report_data |>
  group_by(age_group) |>
  summarise(
    Cases = n(),
    `% of Total` = sprintf("%.1f%%", n() / nrow(report_data) * 100),
    Hospitalized = sum(hospitalized),
    `Hosp Rate` = sprintf("%.1f%%", sum(hospitalized) / n() * 100),
    Deaths = sum(outcome == "Deceased"),
    CFR = sprintf("%.2f%%", sum(outcome == "Deceased") / n() * 100),
    .groups = "drop"
  )

age_summary |>
  gt() |>
  tab_header(title = "Age Group Analysis") |>
  cols_align(align = "right", columns = -age_group) |>
  tab_options(table.font.size = px(12))
```

## Daily Trend

```{r trend-plot, echo=FALSE, fig.width=10, fig.height=4}
daily_data <- report_data |>
  group_by(report_date) |>
  summarise(
    cases = n(),
    hospitalizations = sum(hospitalized),
    .groups = "drop"
  )

p <- ggplot(daily_data, aes(x = report_date)) +
  geom_line(aes(y = cases, color = "Cases"), linewidth = 1) +
  geom_point(aes(y = cases, color = "Cases"), size = 2) +
  geom_line(aes(y = hospitalizations * 5, color = "Hospitalizations (x5)"),
            linewidth = 1, linetype = "dashed") +
  scale_color_manual(values = c("Cases" = "#003366", "Hospitalizations (x5)" = "#DC143C")) +
  labs(
    title = "Daily Case and Hospitalization Trend",
    x = "Date",
    y = "Cases",
    color = NULL
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

ggplotly(p)
```

## Data Quality Assessment

```{r quality-table, echo=FALSE}
quality_summary <- report_data |>
  group_by(data_source) |>
  summarise(
    Records = n(),
    `% of Total` = sprintf("%.1f%%", n() / nrow(report_data) * 100),
    `Lab Confirmed` = sprintf("%.1f%%", mean(lab_confirmed) * 100),
    `Avg Days to Report` = sprintf("%.1f", mean(days_to_report, na.rm = TRUE)),
    .groups = "drop"
  ) |>
  arrange(desc(Records))

quality_summary |>
  gt() |>
  tab_header(
    title = "Data Source Quality Metrics",
    subtitle = "Track SAS migration progress"
  ) |>
  cols_align(align = "right", columns = -data_source) |>
  data_color(
    columns = data_source,
    fn = function(x) {
      case_when(
        x == "Legacy SAS" ~ "#ffcccc",
        x == "Databricks" ~ "#ccffcc",
        TRUE ~ "#ffffff"
      )
    }
  )
```

## Active Outbreaks

```{r outbreaks, echo=FALSE}
active_outbreaks <- outbreaks |>
  filter(status %in% c("Active", "Under Investigation")) |>
  select(outbreak_id, outbreak_type, condition, county, setting, total_cases, hospitalizations, status) |>
  arrange(desc(total_cases))

if (nrow(active_outbreaks) > 0) {
  active_outbreaks |>
    gt() |>
    tab_header(
      title = "Active Outbreak Investigations",
      subtitle = sprintf("%d active investigations", nrow(active_outbreaks))
    ) |>
    data_color(
      columns = status,
      colors = scales::col_factor(
        palette = c("#FF8C00", "#DC143C"),
        domain = c("Under Investigation", "Active")
      )
    )
} else {
  cat("No active outbreaks during this reporting period.")
}
```

## Methodology Notes

This report integrates data from multiple sources:

- **Legacy SAS**: Historical data pipelines (migration in progress)
- **Databricks**: Modern cloud data platform
- **Direct Entry**: Manual case reports
- **Lab Feed**: Automated laboratory results
- **Hospital EHR**: Electronic health record extracts

Data quality metrics help track the transition from legacy SAS workflows to modern Databricks-based pipelines.

---

*Report generated using Posit Quarto. For questions, contact the CDPH Epidemiology Division.*

*This document is intended for public health officials only.*
